Regression Analysis and Prediction

### 2023-10-31 Codes/0_2_quickstat.Rmd
clean up the codes

---
title: "Get crop yield and acreage data: Corn and Soybeans"
---

# Preparation
Read the counties data

```{r }
base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
```

Get the list of states

```{r }
all_states <-
  base_counties$state_name %>%
  unique() %>%
  toupper()
```

Define a function to get QuickStat data

```{r }
get_quickstat_data <- function(year, data_item) {
  desc <-
    tidyUSDA::allDataItem %>%
    grep(pattern = data_item, ., value = TRUE)

  temp_data <-
    tidyUSDA::getQuickstat(
      key = key_get("usda_nass_qs_api"),
      program = "SURVEY",
      data_item = desc,
      geographic_level = "COUNTY",
      state = all_states,
      year = year,
      geometry = FALSE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
    rename("state_code" = "state_fips_code")

  return(temp_data)
}
```

List of items to download

```{r }
item_list <-
  c(
    "CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED"
  )

# tidyUSDA::allDataItem %>%
#     grep(pattern = "PLANTED", ., value = TRUE) %>%
#     grep(pattern = "SOYBEANS", ., value = TRUE) %>%
#     grep(pattern = "IRRIGATED", ., value = TRUE)

# for (i in 1:length(item_list)) {
#   desc <-
#     tidyUSDA::allDataItem %>%
#     grep(pattern = data_item, ., value = TRUE)
#   print(desc)
# }
```

List of year-item combinations

```{r }
par_list <-
  data.table::CJ(
    data_item = item_list,
    year = as.character(1985:2018)
  )
```

@@

# === save the data ===#
saveRDS(nass_data, here("Shared/Data/RawData/nass_data_raw.rds"))
```

@@

```{r process-nass-data}
(
  nass_data_wide <-
    here("Shared/Data/RawData/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not? ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

Quick Summary

```{r }
library(modelsummary)
datasummary(
  crop * ir * (Mean + SD)
  ~ acres + yield,
  data = nass_data_wide
)
```

@@

```{r eval = F}
saveRDS(nass_data_wide, here("Shared/Data/ProcessedData/nass_data.rds"))
```

>>>>

---
title: "Get crop yield and acreage data from USDA-NASS: Corn and Soybeans"
---

# Preparation

```{r }
#--- load the counties ---#
base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))

#--- get the list of states ---#
all_states <-
  base_counties$state_name %>%
  unique() %>%
  toupper()

#--- list of items to download ---#
item_list <-
  c(
    "CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED"
  )

#--- list of year-item combinations ---#
par_list <-
  data.table::CJ(
    data_item = item_list,
    year = as.character(1985:2018)
  )
```

@@

#--- save the data ---#
saveRDS(nass_data, here("Data/data-raw/nass_data_raw.rds"))
```

@@

```{r process-nass-data}
(
  nass_data_wide <-
    here("Data/data-raw/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not? ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

@@

```{r eval = F}
saveRDS(nass_data_wide, here("Data/data-processed/nass_data.rds"))
```

### 2023-10-31 Codes/0_3_get_sat.Rmd
clean up the codes

---
title: "Assign saturated thickness value "
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi=400,
  fig.width=7.5,
  fig.height=5,
  out.width="750px",
  out.height="500px"
)
```

@@

```{r }
sat_temp <- raster("/Volumes/Elements/Database/SaturatedThickness/SatThick_19851.tif")
```

@@

```{r }
hpa_counties <- 
  here("Shared/Data/ProcessedData/base_counties.rds") %>% 
  readRDS() %>% 
  .[in_hpa == 1, ] %>% 
  st_as_sf() %>% 
  st_transform(projection(sat_temp)) 
```

@@

```{r }
get_sat <- function(y) {
  # y <- 2000
  sat_temp <- 
    paste0("/Volumes/Elements/Database/SaturatedThickness/SatThick_", y, "1.tif") %>% 
    raster()

@@

```{r extract-sat, eval = F}
sat_all <- mclapply(1985:2016, get_sat, mc.cores = 12) %>%
  rbindlist() %>%
  .[sat < 0, sat := 0] %>% 
  .[, sat := sat * 3.28084] # meter to feet
```

@@

```{r save, eval = F}
saveRDS(sat_all, here("Shared/Data/ProcessedData/sat_all.rds"))
```

# Quick visualization

```{r }
sat_all <- 
  here("Shared/Data/ProcessedData/sat_all.rds") %>% 
  readRDS()
```

Saturate thickness history by county

```{r }
ggplot(sat_all) +
  geom_line(aes(y = sat, x = year, group = sc_code))
```

Histogram

```{r }
ggplot(sat_all) +
  geom_histogram(aes(x = sat))
```

>>>>

---
title: "Assign saturated thickness values"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

@@

```{r }
sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
```

@@

```{r }
hpa_counties <-
  here("Data/data-processed/base_counties.rds") %>%
  readRDS() %>%
  .[in_hpa == 1, ] %>%
  st_as_sf() %>%
  st_transform(projection(sat_temp))
```

@@

```{r }
get_sat <- function(y) {
  # y <- 2000
  sat_temp <-
    paste0("Data/data-raw/saturated_thickness/SatThick_", y, "1.tif") %>%
    raster()

@@

```{r extract-sat, eval = F}
sat_all <-
  mclapply(1985:2016, get_sat, mc.cores = 12) %>%
  rbindlist() %>%
  .[sat < 0, sat := 0]
```

@@

```{r save, eval = F}
saveRDS(sat_all, here("Data/data-processed/sat_all.rds"))
```

### 2023-10-31 Codes/0_6_merge_datasets.rmd
clean up the codes

```{r}
base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))

#' NDMC Drought index
drought_county <- readRDS(here("Shared/Data/ProcessedData/DI_NDMC.rds"))

#' SPEI
# spei_data <- readRDS(here("Shared/Data/ProcessedData/spei_data.rds"))

#' Weather
weather_data <- readRDS(here("Shared/Data/ProcessedData", "summrized_gridMET.rds"))

#' Saturated thickness
sat_all <- readRDS(here("Shared/Data/ProcessedData/sat_all.rds"))

#' Yield, acreage data
nass_data <- readRDS(here("Shared/Data/ProcessedData/nass_data.rds"))

#* SSURGO

ssurgo_soil <- readRDS(here("Shared/Data/ProcessedData/ssurgo_soil.rds"))
```

@@

```{r}
merged_data <-
  base_counties[nass_data, on = c("sc_code")] %>%
  weather_data[., on = c("sc_code", "year")] %>%
  sat_all[., on = c("sc_code", "year")] %>%
  drought_county[., on = c("sc_code", "year")] %>%
  .[order(sc_code, year), ] %>%
  ssurgo_soil[, on = "sc_code"]

# merged_data[ir == "nir", state] %>% unique()
# merged_data[ir == "ir", state] %>% unique()
```

@@

```{r}
final_data <-
  merged_data %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  # === if not irrigated, set sat = 0 ===#
  .[ir == "nir", sat := 0] %>%
  # === if sat == 0, but indicated irrigated ===#
  # add 0.1 to distinguish dryland production, where
  # sat is set at 0 above
  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
  # === drop if yield or acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  # === if sat is missing, drop ===#
  #' just 2017 and 2018 data because saturated thickness data
  #' is observed only until 2016
  .[!is.na(sat), ] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_code] %>%
  # === keep only if more than 2 observations are available ===#
  .[num_obs >= 10, ] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ]
```

@@

```{r}
saveRDS(final_data, here("Shared/Data/ProcessedData/final_data.rds"))
```

>>>>

## Read the datasets

```{r}
base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))

#' Weather
weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))

#' Saturated thickness
sat_all <- readRDS(here("Data/data-processed/sat_all.rds"))

#' Yield, acreage data
nass_data <- readRDS(here("Data/data-processed/nass_data.rds"))

#* SSURGO
ssurgo_soil <- readRDS(here("Data/data-processed/ssurgo_soil.rds"))
```

@@

```{r}
merged_data <-
  base_counties[nass_data, on = c("sc_code")] %>%
  weather_data[., on = c("sc_code", "year")] %>%
  sat_all[., on = c("sc_code", "year")] %>%
  .[order(sc_code, year), ] %>%
  ssurgo_soil[, on = "sc_code"]
```

@@

```{r}
final_data <-
  merged_data %>%
  #--- yield unit conversion (bu/acre to ton/ha)---#
  .[, yield := yield * 12.553 / 200] %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  # === if not irrigated, set sat = 0 ===#
  .[ir == "nir", sat := 0] %>%
  # === if sat == 0, but indicated irrigated ===#
  # add 0.1 to distinguish dryland production, where
  # sat is set at 0 above
  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
  # === drop if yield or acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  # === if sat is missing, drop ===#
  #' just 2017 and 2018 data because saturated thickness data
  #' is observed only until 2016
  .[!is.na(sat), ] %>%
  #--- indicator of whether dryland (rainfed) production or not ---#
  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
  #--- sc_code and dry_or_not (used for FE regression) ---#
  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
  #--- state-year combination ---#
  .[, state_year := paste0(state_name, "_", year)] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_dry] %>%
  # .[, num_obs := .N, by = sc_code] %>%
  # === keep only if more than 10 observations are available ===#
  .[num_obs >= 10, ] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ]
```

@@

```{r}
saveRDS(final_data, here("Data/data-processed/final_data.rds"))
```

### 2023-10-31 Codes/1_Analysis.rmd
clean up the codes

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    #--- saturated thickness has to be at least 9 meters ---#
    sat_thld_m = 9,
    #--- at least 75% of total county area has to be overlapped with HPA ---#
    ir_share_thld = 0.75,
    #--- drop obs with extreme balance ---#
    balance_thld = c(-1200, 200),
    #--- number of saturated thickness categories ---#
    sat_cat_num = 3,
    #--- select states of focus ---#
    # no observations are dropped with the current set of states
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop")

@@

saveRDS(main_analysis, "Shared/Results/main_analysis.rds")
```

@@

saveRDS(all_results, "Shared/Results/all_results.rds")
```

>>>>

```{r}
#--- read data ---#
final_data <-
  here("Data/data-processed/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    #--- saturated thickness has to be at least 9 meters ---#
    sat_thld_m = 9,
    #--- at least 75% of total county area has to be overlapped with HPA ---#
    ir_share_thld = 0.75,
    #--- drop obs with extreme balance ---#
    balance_thld = c(-200, 1200),
    #--- number of saturated thickness categories ---#
    sat_cat_num = 3,
    #--- select states of focus ---#
    # no observations are dropped with the current set of states
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop")

@@

saveRDS(main_analysis, "Results/main_analysis.rds")
```

@@

saveRDS(all_results, "Results/all_results.rds")
```
